Fish do not need Visas: the transboundary nature of the world’s exploited marine fish species

### 2019-12-23 Manuscript_Draft.Rmd
x
Updates in results and discussion

## Methods summary (250/250 words)

@@

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79.7** million in fishing revenue, equivalent to **49.8**% of the total revenue of the same period. These values are higher than previously estimated [@Teh] and higlight the importance of these stocks at the global level.

Transboundary species are particularly important at the regional and national level. In terms of catch and revenue, Eastern Asia and Latin America and the Caribbean are the higher zones (Fig 1B). Particular for counties in North America (Bermuda, Canada, Greenland, and the US), a moderate amount of catch (3.4 thousand tons) are worth 11.7 billion USD, highlighting the importance of transboundary species to the economy of the region. The EEZs along Western Eurpe, in turn, are home of **X** species responding to the joint fisheries management policy established by the European Union. In here, quotas are allocated by species and regulatory area [@Melnychuk:2017gh], making this region **unique** in terms of joint management of a high numer of shared stocks.

At the country level, transboundary species are most valuable for the United States (USD 9,605 million), Russia (USD 6,361 million), China (USD 6,279 million) and Peru (USD 5,211). On the other hand, EEZs like that of Portugal, the Mediterranean side of Spain, Senegal, and the UK are rich in transboundary species (n > 110), but these do not represent a large amount of revenue, comparativeley. China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018], is the only country categorized high in both number of transboundary species (n = 119) and revenue. For the US, 16 species make up 90% of the revenue from transboudary species, with American scallop (*Placopecten magellanicus*) shared with Canada in Georges Bank [@https://elischolar.library.yale.edu/cgi/viewcontent.cgi?article=1008&context=fes-pubs] contributing 41.8% or the revenue. Russia shares the largest fished species in the world, Alaskan polllock (*Theragra chalcogramma*) [@SOFIA2018], with both Japan and the US. For Russia, this fishery is responsible for 50% of the revenue from transboundary species with other eight species comprising the 90% of revenue. In the case of Peru, Peruvian anchoveta (*Engraulis ringens*) and jumbo squid (*Dosidicus gigas*), both shared with Chile, represent almost 90% of the revenue from transboundary species, with anchoveta accounting for 79%.

These results hilights the importance of transboundary species not only at the global and regional levels, but also at the country level.

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="A) Number of transboundary species per EEZ and they contribution to countries fishing revenue. The number of species is represented in the EEZ while the revenue is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species."}

@@

![**A) Number of transboundary species per EEZ and they contribution to countries fishing revenue. The number of species is represented in the EEZ while the revenue is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species.**](./Figures/Fig1.png)

@@

![**Number of transboundary and discrete species by country and continent**](./Figures/Fig2.png)

Empirical analysis suggest that cooperation will have better outcomes for shared fisheries management. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*)[@Diekert:2010gp], and between Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw]. However, few evidence exists today regarding the true efficiency of transboundary fisheries plans. Indeed, there are some successful bilateral management plans for species such as Pacific Halibut between Canada and the US as well [@REference?] and multilateral treaties like the Pacific Nauru Agreement that manages 30% of global skipjack tuna in the South Pacific Islands [@Aqorau:2018bh]. However, the effectiveness of Regional Fisheries Management Organizations, responsible for managing high migratory stocks in areas beyond national jurisdiction, has been previously questioned [@CullisSuzuki:2010fi]. Countries that are today harvesting species at rates less than 50% of the historical maximum catch (Category C) should be paying extra care in these species (Fig. 2). While these catch trends do not reflect stock status [@@Branch:2011br], they could reassemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. The rebuilding of a transboundary species could be strongly impacted by uncoordinated efforts of countries harvesting the same species but with different objectives [@Ref].

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@REF]. Maybe not surprisingly, highly migratory pelagic-oceanic species such as Wahoo (*Acanthocybium solandri*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are among the more shared resources in the world with a median of 40 nations sharing these resources. These species are very important in terms of economic value as they (global) average revenue exceeds USDXXX. The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even if a population swims between the borders of two nations, their larvae could be supplying populations thousands of kilometers away [@Ramesh:2019va]. Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf].

>>>>

<!-- ## Methods summary (250/250 words) -->

@@

Overall, we estimated **633** transboundary species in the world, more than double the number previously estimated [@Teh:2015gd], accounting for **65.4**% of the species analyzed (Fig. 1A). Between 2005 and 2014, national fleets targeting shared stocks within EEZs landed an average of **69.2** million tonnes, representing **64**% of the EEZ-global catch, and generating USD **79.7** million in fishing revenue. This is equivalent to **49.8**% of the global revenue of the same period. These values are higher than previously estimated reinforcing the importance of these stocks at the global level [@Teh].

Transboundary species are particulary important for Northern America (mean = 3,901 $\pm$ 4,497 million USD) and Eastern Asia (mean = 2,827 $\pm$ = 2,667 million USD), significantly differing from the rest in tn terms of revenue (one-way ANOVA; *F*(16,157) = 4.91, *p* < 0.001, $\alpha$ = 0.05) (Fig 1B). The large variation in revenue from these regions is driven by China (USD 6,279 million), in E. Asia and the US (USD 9,605 million) in N. America who contribue 55% and 83% of the region's revenue, respectiveley. Russia (USD 6,361 million), Peru (USD 5,211) and Japan (USD 3,369) complete the top five countries with more revenue from transboundary species.

Catches, however, are only significantly different in E. Asia (one-way ANOVA; *F*(16,155) = 3.341, *p* < 0.001, $\alpha$ = 0.05), with China, the world's top fish producer with 15.3 million tons landed in 2016 [@SOFIA2018], getting around 1/3 of its capture from transboundary species. For the US, 16 species make up 90% of the revenue from transboudary species, with American scallop (*Placopecten magellanicus*) shared with Canada in Georges Bank [@https://elischolar.library.yale.edu/cgi/viewcontent.cgi?article=1008&context=fes-pubs] contributing 41.8% or the revenue. Russia shares the largest fished species in the world, Alaskan polllock (*Theragra chalcogramma*) [@SOFIA2018], with both Japan and the US. Rhis fishery is responsible for 50% of the Russian revenue from transboundary species, with other eight species comprising the 90% of revenue. In the case of Peru, Peruvian anchoveta (*Engraulis ringens*) and jumbo squid (*Dosidicus gigas*), both shared with Chile, represent almost 90% of the revenue from transboundary species, with anchoveta accounting for 79%.

```{r Call_Map, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="A) Number of transboundary species per EEZ and they contribution to countries fishing revenue. The number of species is represented in the EEZ while the revenue is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species.Points = mean +- sd."}

@@

![**A) Number of transboundary species per EEZ and they contribution to countries fishing revenue. The number of species is represented in the EEZ while the revenue is represented in the land polygon. B) Regional revenue (left) and catch (right) contribution of tranbsoundary species. Points = mean +- sd.**](./Figures/Fig1.png)

@@

There is no significant difference in the exploitaion rate of transboundary species between continents (one-way ANOVA *F*(5,443), *f* = 1.24, *p* = 0.289, $\alpha$ = 0.05) as most species are fished over 50% of their historical maximum peak (Category C) in all continents, but Oceania (Fig.2). However, the number of species exploited by each country in each category significantly differs from each other (one-way ANOVA *F*(5,442), *f* = 322.9, *p* < 0.001, $\alpha$ = 0.05). Overall, most transboundary species are fished over 50% of their once maximum peak (Category C) suggesting that their current exploitation rate is over the maximum sustainable yield [@Kleisner:2011wn].

Empirical analysis suggest that cooperation will have better outcomes for shared fisheries management. Examples vary between regions and species including Mexico and the United States over Pacific anchovies [@CisnerosMontemayor:2020kv], Norway and Russia over Atlantic cod (*Gadus morhua*)[@Diekert:2010gp], and between Namibia and South Africa over hake (*Merluccius spp*) [@Sumaila:2003vw].

Indeed, there are some successful bilateral management plans for species such as Pacific Halibut between Canada and the US as well [@REference?] and multilateral treaties like the Pacific Nauru Agreement that manages 30% of global skipjack tuna in the South Pacific Islands [@Aqorau:2018bh]. However, the effectiveness of Regional Fisheries Management Organizations, responsible for managing high migratory stocks in areas beyond national jurisdiction, has been previously questioned [@CullisSuzuki:2010fi]. Countries that are today harvesting species at rates less than 50% of the historical maximum catch (Category C) should be paying extra care in these species (Fig. 2). While these catch trends do not reflect stock status [@Branch:2011br], they could reassemble fisheries that are over exploited or collapsed [@Kleisner:2011wn]. The rebuilding of a transboundary species could be strongly impacted by uncoordinated efforts of countries harvesting the same species but with different objectives [@Ref].

```{r Call_Circ_Country, eval = F, fig.width=6, fig.height=4, echo=FALSE, fig.pos= "H", fig.cap="Number of transboundary species by catch category, country, and continent. Category A, fishing between 10%-50% of max catch peak after a drop larger than 10% of max peak; Category B, fishing over 50% of max catch peak; Category C, fishing is between 10-50% of max catch peak or less than 10%"}

img <- readPNG(here("./Figures/Fig2.png"))
grid.raster(img)
```

![**Number of transboundary species by catch category, country, and continent. Category A, fishing between 10%-50% of max catch peak after a drop larger than 10% of max peak; Category B, fishing over 50% of max catch peak; Category C, fishing is between 10-50% of max catch peak or less than 10%**](./Figures/Fig2.png)

Species can have a cosmopolitan distribution, that is, a broader distribution, or be limited to a specific region, known as endemic [@REF]. Not surprisingly, highly migratory pelagic-oceanic species such as skipjack tuna (*Katsuwonus pelamis*), Common thresher shark (*Alopias vulpinus*), and tunas (*Thunnus sp.*) are among the more shared resources in the world with a median of 40 nations sharing these resources (Fig 3.). These species are very important in terms of economic value as they (global) average revenue exceeds USDXXX.

The median for species of all other ecosystem preferences is close to, or less than, 20, as many of these might be cosmopolitan but not necessarily highly migratory (e.g. *Solea solea*, *Panulirus argus*) (Fig. 3). However, less distributed species must not be ignored. The world of fisheries is highly connected trough larvae dispersal [@Ramesh:2019va] and so, even if a population swims between the borders of two nations, their larvae could be supplying populations thousands of kilometers away [@Ramesh:2019va].

Moreover, many coastal communities, specially in developing regions, and highly dependent on fishing local fish (e.g. reef fish) for food security and improved livelihoods [@CisnerosMontemayor:2016gq; Wabnitz:2018gf].
